---
title: "Homework 3,.1"
author: "Drew Remmenga"
date: "2024-07-14"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\pagebreak
Exercise 4.4 (a)
Categorical variables: sex, future plans, citizenship
Numerical variables: age at graduation, time to complete degree, class rank, grade point average
Nominal: Sex, citizenship
Ordinal: future plans
Discrete: Class rank
Cointinuous: Age at graduation, grade point average
\clearpage
Exercise 4.4 (b)
Ratio scale: age at graduation, time to complete degree
Interval scale: class rank, grade point average
\clearpage
Exercise 4.7 (a)
```{r}
.146*6276+.087*5212+.386*6928+6127*.494+.465*4094+.581*6397+.346*2417
```
```{r}
library(qcc)
defect <- c(.146*6276,.087*5212,.386*6928,6127*.494,.465*4094,.581*6397,.346*2417)
names(defect) <- c("Physical Sciences","Engineering","Life Sciences","Social Sciences","Humanitites","Education","Professional Fields")
pareto.chart(defect,xab="Categories",ylab="frequency")
```
\clearpage
Exercise 4.7 (b)
Most degrees are awarded in education with very few in engineering. 