---
title: "Homework 5.2"
author: "Drew Remmenga"
date: "2024-07-24"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Exercise 6.11 
(a)

$\bar{x} \pm 1.645*\sigma/\sqrt{n}$ 95% interval.$\\$
(b)
```{r}
30+1.645*10/sqrt(100)
30-1.645*10/sqrt(100)
```
(c)
95% $\\$
Exercise 6.13 (a)
95 $\\$
(b)
The width would decrease.
$\\$
Exercise 6.14 (a)
```{r}
16.3+1.282*6/sqrt(25)
16.3-1.282*6/sqrt(25)
16.3+1.645*6/sqrt(25)
16.3-1.645*6/sqrt(25)
16.3+2.576*6/sqrt(25)
16.3-2.576*6/sqrt(25)
``` 
(b)
The widths would decrease. $\\$
Exercise 6.15 (a)
```{r}
110.5+1.960*3/sqrt(10)
110.5-1.960*3/sqrt(10)

```
We are 95% confidenct that the true mean lies in this range. Thus the specifications are met 95% of the time in this interval.
$\\$
(b)
```{r}
110.5-1.645*3/sqrt(10)
```
We are 95% confident that the value of the mean voltage will be above this value. Thus the specifications are met.
$\\$
(c)
```{r}
110.5+1.645*3/sqrt(10)
```
We can tell the speicifcation is met at least 95% of the time since this confidence interval is bleow the given value. $\\$
Exercise 6.16
50%
```{r}
1-.5^(10)
```

