---
title: "Homework 5.3"
author: "Drew Remmenga"
date: "2024-07-24"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Exercise 6.17 (a)
$H_{0} P=.54 \\$
$H_{1} P \neq .54 \\$
(b)

$H_{0} P\geq .05 \\$
$H_{1} P < .05 \\$
(c)

$H_{0} P\geq .40 \\$
$H_{1} P < .40 \\$
(d)

$H_{0} P_{gatorade}\leq P_{All sports} \\$
$H_{1} P_{gatorade} > P_{all sports} \\$
Exercise 6.19
a) The null hypothesis should be - The chemical compound used as a food additive is a suspected carcinogen and is not consumed in the safe amount. Not detecting a cancer cause is a serious threat.$\\$

(b) The null hypothesis should be - The new drug is not bioequivalent to the original drug, because that is a critical factor for drug approval by the FDA. If it is incorrectly concluded that the drug is bioequivalent, then a potentially hazardous drug may come out in the market for general use. $\\$
(c) 
It is not biologicallyt equivalent should be set up as the null hypothesis as people could get sick otherwise.$\\$
(d)
It is effective should be set up as the null hypothesis because believing it is effective is important for preventing drought. 
$\\$
Exercise 6.20 (a)
50% 50%
$\\$
(b)
.25% .75% $\\$
Exercise 6.27 (a)
H0 =10 $\\$
H1< 10 $\\$
(b)
```{r}
z<-(8.7-10)/(2/sqrt(15))
pnorm(z,0,1)
```
refject null hypothesis is there is enough evidenc eto support the claim at .05; $\\$
(C)
```{r}
.1039
1-.1039
```
Exercise 6.30 (a)
```{r}
1-.95^20
```
(b)
Whenever there is multiple testing to be done there is increased probability of a type I error. 