---
title: "Homework 8.3"
author: "Drew Remmenga"
date: "2024-08-11"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Exercise 9.20 (a) H0: p1=9/16,p2=3/16,p3=3/16,p4=1/16 $\\$
H1: at least mone proportion is different from the stated version. $\\$
(b)
```{r}
chi=1.269
df=3
p=.6895
p
chi
```
We fail to reject the null hypothesis at the .05 level. $\\$
Exercise 9.22 (a)
```{r}
N=678+227+56+28+8+14
observed=c(678,227,56,28,8,14)
expected=c(676.9,224.9,63.1,16.4,3.7,1.1)
chi=(observed-expected)^2/expected
chi=21.021
p=.001
chi
p
```
Since p is low we reject the null hypothesis.
(b)
```{r}
expected=c(676.9,224.9,85.9,32.8,12.3)
chi=(observed-expected)^2/expected
chi=17.239
p=.002
chi
p
```
Since p is low we reject the null hypothesis.
Exercise 9.28 (a) Chi squared of rindependence since we want to find out if they are independent. $\\$
(b)
H0: Returned are dpendent on citites. 
H1: returned depend on type of city.
(c)
```{r}
chi=4.41
df=3
```
Do not reject null hypothesis. $\\$ 
Exercise 9.30 (a)
```{r}
sick=c(31,60)
notsick=c(77,48)
treatment=c(31,60)
placebo=c(60,48)

```
(b)
H0:We nare testing being sick is independent of treatment.
H1: Being sick is not dependent on receiving treatment
(c)
```{r}
chi=15.9698
df=1
```
Since p is low we reject H0 and conclude dramimine is an effective drug. 